Supporting Information: Improving cropland maps through tight integration of human and machine intelligence

### 2020-07-24 inst/manuscript/si.Rmd
Substantial updates to analyses and organization. Manuscript, figures notebook, and analyses moved to inst/manuscript. Added scripts to data-raw/ for processing inputs into figures, headed to instances, etc (much of this is moving code out of analyses* notebooks).

---
title: "Supporting Information"

author:

- name: \*Provisional author list\*
  affilnum: '1'

- name: Lyndon D. Estes
  affilnum: '1'
  email: lestes@clarku.edu

- name: Su Ye
  affilnum: '1'

- name: Lei Song
  affilnum: '1'

- name: Ron Eastman
  affilnum: '1'

- name: Sitian Xiong
  affilnum: '1'

- name: Tammy Woodard
  affilnum: '1'

- name: Boka Luo
  affilnum: '1'

- name: Dennis McRitchie
  affilnum: '2'

- name: Ryan Avery
  affilnum: '3'

- name: Kelly Caylor
  affilnum: '3'

- name: Stephanie Debats
  affilnum: '4'

- name: SpatialCollective
  affilnum: '5'

- name: Meridia
  affilnum: '6'

- name: Azavea
  affilnum: '7'

affiliation:

- affilnum: 1
  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA 01610 USA

- affilnum: 2
  affil: Dennis and Sons, Tucson, AZ, USA

- affilnum: 3
  affil: UCSB

- affilnum: 4
  affil: Uber

- affilnum: 5
  affil: SpatialCollective

- affilnum: 5
  affil: Meridia

- affilnum: 5
  affil: Azavea

output:

  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: yes
    template: manuscript.latex
    includes:
      in_header: header.tex

@@

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: false
spacing: doublespacing
footerdate: yes
abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

@@

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(knitr)
library(citr)
library(tidyverse)
library(sf)

@@

\singlespace

\bleft
# Methods
## Assessing quality of composites
The rubric presented in Table 1 was used to assess the quality of the seasonal image composites. The imagery was evaluated by examining their Raster Foundry overlays within a `labeller` instance set up for the purpose.

@@

@@

Since `r round((wtA * 1 + wtB * 0) / (wtA + wtB), 3)` > 0.5, we label the particular pixel a field.

@@

Where p is the estimated probability (or mapped class accuracy), and d is the size of the margin of error (1/2 the confidence interval). We selected a d value of 0.03 and assumed that the user's accuracy of the field class would be 0.75 and that of the non-cropland class would be 0.8, returning sample sizes of 800 and 683, respectively. The distribution of the resulting map reference sample is shown in Figure 1.

```{r, echo = FALSE, fig.align='center', out.width="80%", message=FALSE, warning=FALSE, fig.cap="Distribution of the selected map reference sample for the cropland and non-cropland class. "}
africa <- st_read(
  system.file("extdata/africa.geojson", package = "activemapper"), quiet = TRUE,
)
aois <- st_read(system.file("extdata/aois.geojson", package = "activemapper"), 
                quiet = TRUE)
africa_cont <- africa %>% st_union %>% st_sf
ghana <- africa %>% filter(name == "Ghana") %>% select(geometry)

ref_sample <- readr::read_csv(
  here::here("external/reference/map_reference_sample.csv")
)
ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)

p <- ggplot() + 
  # geom_sf(data = ghana) +
  geom_sf(data = aois) + 
  geom_sf(data = ref_sample, aes(color = factor(class)), size = 0.1) +
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Non-cropland", "Cropland"), name = "") + 
  theme_light()
p
```

@@

@@

# References
\singlespace

>>>>

---
title: "Supporting Information: Improving cropland maps through tight integration of human and machine intelligence"

output:
  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: yes
    template: manuscript.latex
    includes:
      in_header: header.tex

@@

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: false
spacing: doublespacing
footerdate: yes
# abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

@@

<!-- Provides custom captions for figures in supplement -->
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(knitr)
library(citr)
library(tidyverse)
library(sf)
library(activemapper)

@@

\beginsupplement
\singlespace

\bleft

# Methods
## Assessing quality of composites
The rubric presented in Table 1 was used to assess the quality of the seasonal image composites. The imagery was evaluated by examining their Raster Foundry overlays within a `labeller` instance set up for the purpose.

@@

After creating the consensus label, the degree of confidence in the resulting label value can be expressed in terms of Bayesian risk. Risk for a given training sample site *i* is calculated as:

\begin{equation}
\mathrm{R_i=\frac{1}{n}\sum_{j=1}^{n} C_j(1 - L_j) + (1 - C_j)L_j}
\end{equation}

Where *C* is the consensus probability that a given pixel *j* is a field ($\mathrm{P(\theta = field|D)}$), and *L* is the consensus label (i.e. non-field if *C* < 0.5, field if *C* > 0.5) for that pixel. *R* is thus the average risk across all pixels at a given site, and the slope of the risk varies depending on whether *L* is a field or not a field (Figure \ref{fig:riskcurve}). The closer to 0 the lower the risk that the *L* is mislabelled, while values approaching 1 indicate increasing risk of mislabelling.

```{r riskcurve, echo = FALSE, fig.align='center', out.width="70%", message=FALSE, warning=FALSE, fig.cap="Bayesian risk values (Y-axis) for consensus values (X axis) ranging from 0 to 1 (0 indicates no consensus that a pixel falls into the field class, 1 means complete consensus) for field and non-field consensus labels."}
knitr::include_graphics('figures/si_label_risk.png')
```

@@

Since `r rj <- round((wtA * 1 + wtB * 0) / (wtA + wtB), 3); rj` > 0.5, we label the particular pixel a field.

Using the relevant part of equation 15, T=the corresponding risk associated with this particular pixel's label $\mathrm{R_j}$ is thus `r rj * (1 - 1) + (1 - rj) * 1`.

@@

Where p is the estimated probability (or mapped class accuracy), and d is the size of the margin of error (1/2 the confidence interval). We selected a d value of 0.03 and assumed that the user's accuracy of the field class would be 0.75 and that of the non-cropland class would be 0.8, returning sample sizes of 800 and 683, respectively. The distribution of the resulting map reference sample is shown in Figure \ref{fig:refsample}.

```{r refsample, echo = FALSE, fig.align='center', out.width="80%", message=FALSE, warning=FALSE, fig.cap="Distribution of the selected map reference sample for the cropland and non-cropland class. "}
knitr::include_graphics('figures/si_map_reference_sample.png')
```

@@

## Active learning process
### Labelling
The distributions of the initial 500 sites used to train the RandomForest model for each AOI are shown in Figure \ref{fig:trainval}. In AOI 3, the initial active learning cycle resulted in low accuracy because the northern part of the AOI shows low contrast between fields and the surrounding vegetation in the dry season. Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
```{r trainval, echo = FALSE, out.width="80%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected within AOI 3 to retrain a second active learning run for this region, as well as the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_training_validation_pts.png')
```

The distribution of data collection effort and scores received at training reference sites by labelling team (Figure \ref{fig:trainqual}).

```{r trainqual, echo = FALSE, out.width="60%", fig.cap="The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller. Labellers' identities are anonymized.", fig.align='center', message=FALSE, results='hold'}
knitr::include_graphics('figures/si_training_quality_n.png')
```

## Experiments
### Label quality

@@

\clearpage
# References
\singlespace
